### 2021-05-12 Latex/sections/results.tex

%!TEX root = ../main.tex
\section{Results}

\subsection{Simulated Results}

The plane-based registration is applied to the simulated dataset with noisy pose and range measurements (cf. Figure~\ref{fig:simulatedDatasets}) without further processing.
Assuming this represents a coarsely pre-registered 3D point cloud the distances to the ground truth were evaluated before and after the plane-based registration. 
Figure~\ref{fig:simulatedEvaluation} shows the different point-to-point distances.

\begin{figure*}
 	\centering
 	\begin{minipage}[c]{0.49\textwidth}
 		\centering
		\includegraphics[width=\textwidth]{./images/uncorr_bottom_pose}\\
		\includegraphics[width=\textwidth]{./images/uncorr_side_view}\\
  		\includegraphics[width=\textwidth]{./images/uncorr_hist}
  	\end{minipage}\hfill
  	\begin{minipage}[c]{0.49\textwidth}
  		\centering
		\includegraphics[width=\textwidth]{./images/corr_bottom_pose}\\
		\includegraphics[width=\textwidth]{./images/corr_side_view}\\
  		\includegraphics[width=\textwidth]{./images/corr_hist}
  	\end{minipage} 	
 	\caption{Evaluation of point distances before (left) and after (right) the plane based registration on a simulated dataset. A maximal distance of \SI{2}{m} is set, such that all points that display a higher distance value are excluded from analysis. Further the histogram joins all values with a distance greater than \SI{1}{m} into the last bucket. The top two column shows a heat map of distances while the bottom shows the corresponding histogram. The color mapping is equivalent in both. An animation of the matching process is given at~\href{https://youtu.be/0ps5Pg4qo4E}{here}}
 	\label{fig:simulatedEvaluation}
\end{figure*}

Before the registration the corridor is only represented acceptably in the front part. 
The further into the corridor, i.e. the longer the robot accumulates error, the more imprecise the date becomes. 
Finally we see that a large portion of the data even exceeds either the \SI{1}{\meter} error, contributing to the large spike in the histogram or even the \SI{2}{\meter} error limit thus, being cut from the representation. 
After registration, we see that, qualitatively, the ideal corridor was nearly restored from the noisy data. 
In particular a very large portion (95\%) have less than \SI{17}{\centi\meter} distances.
Further the square and straight shape of the corridor is restored well and especially the large spike at errors of greater than \SI{1}{m} is remove. 
Any such errors tend to occur at the back and the front of the corridor where the measured range is the largest, hence has the largest contribution of the range error.

\subsection{Floating Sphere Results}

Figure~\ref{fig:cylon-corrected} shows the results obtained after employing the plane base registration on the dataset acquired by the floating sphere.
\begin{figure*}
	\centering
	\includegraphics[width=0.49\textwidth]{./images/cylon_uncorr_corner}\hfill
	\includegraphics[width=0.49\textwidth]{./images/cylon_corr_corner}\\
	\includegraphics[width=0.49\textwidth]{./images/cylon_uncorr_top}\hfill
	\includegraphics[width=0.49\textwidth]{./images/cylon_corr_top}
	\caption{The point cloud acquired by the floating sphere before (left) and after (right) applying the plane based registration. View from the interior that shows the used planes (top) and a birds-eye view (bottom). An animation of the registration process is given at \href{https://youtu.be/8XdIUN_9VpY}{here}.}
	\label{fig:cylon-corrected}
\end{figure*}
After the registration the walls of the room are significantly more prominent in the point-cloud. 
In particular the noise in the top left corner of the top view (cf. figure~\ref{fig:cylon-corrected}) has been visibly reduced.
Further, the noise around the walls is notably reduced as the points are moved on the plane.

However, some imprecisions remain. 
At the same location a few points cannot be matched to the plane as they are clustered with some other points that are bound to a different pose. 
Hence this error stems from the imprecise interpolation in between scans. 
\todo[inline]{More comments here}

\subsection{RADLER Results}

\subsection{Experimental Results REMOVE THIS SECTION}

For the experimentally acquired datasets a similar procedure is followed. 
The main difference being, that the scan was pre-registered using a point-to-point based method as described in section~\ref{sec:experimentalSetup}. 
Here also the point distances to the ground truth (in this case the terrestrially acquired 3D laser scan) is evaluated before and after the plane-based registration process.

\begin{figure}
	\centering
	\begin{minipage}[c]{0.25\textwidth}
		\centering	
		\includegraphics[width = \textwidth]{./images/distances_fire}\\
		\includegraphics[width = \textwidth]{./images/histogram_fire}
  	\end{minipage}\hfill
  	\begin{minipage}[c]{0.25\textwidth}
  	\end{minipage}

	\todo[inline]{Evaluate Plane matching on firefighter school data; histogram after etc}
	\caption{Evaluation of point distances before (left) and after (right) the plane based registration on a experimentally acquired dataset. A maximal distance of \SI{5}{m} is set, such that all points that display a higher distance value are excluded. The left column shows a heat map of distances while the right shows the corresponding histogram. The color mapping is equivalent in both. An animation of the matching process can be seen \href{todo}{here}}
	\label{fig:experimentalEvaluation}
\end{figure}