### 2021-05-12 Latex/sections/simulation.tex

To quantify the quality of the proposed registration method, it is tested on simulated as well as on real world data.

%!TEX root = ../main.tex
\section{Simulation}

For the simulated data we implemented a noisy world-robot-sensor model.  
We are simulating a LIVOX-Mid100~\cite{LivoxMid40-100} laser scanner with customizable noise level on the range measurements inside a robot with different motion capabilities, subject to noise in its pose estimation. This yields scan results similar to the ones obtained in the real world.

\subsection{Range Noise Model}

Most real laser scanners are subjected to range measurement noise that is proportional to the measured range. 
Therefore the simulated sensors are as well. 
To achieve this we sample a noise percentage $n_P$ from some normal distribution $\mathcal{N}(\mu,\sigma^2)$. 
The resulting range $r$ given the true range $r_t$ is then
\begin{align}
	r = r_t(1+n_P)
\end{align}
For each measurement ray the noise percentage is sampled independently.

\subsection{Pose Estimation Noise Model}

For the pose measurement one cannot simply add white noise to the current pose estimate as this does not capture the integration error which is common among inertial measurement units (IMU). 
Therefore we assume a disturbance torque about the two axes that lie in the ground plane. 
This is equivalent to a case, where a slightly shifted ground plane or a unbalanced locomotion system of the robot is present.

Accordingly, a spherical robot that is assumed to be translating exactly along one axis by rotation exactly about one other axis at an angular velocity $\omega$  experiences random disturbance torques at each time instant that accelerate the sphere's rotation about the respective axes. 
This additional motion is determined by integrating the disturbance torques which are sampled from some normal distribution $n_\phi, n_\psi \in \mathcal{N}(\mu,\sigma^2)$ each.  
The pose update in the discrete simulation is therefore given by: 
\begin{align}
	\begin{bmatrix}\varphi\\\vartheta\\\psi\\x\\y\\z\end{bmatrix}_{k+1} = 
	\begin{bmatrix}\varphi\\\vartheta\\\psi\\x\\y\\z\end{bmatrix}_{k} 
	+ \begin{bmatrix}1\\0\\0\\R\\0\\0\end{bmatrix}\omega \Delta t  + \Delta t \underbrace{\sum_{i = 0}^k \begin{bmatrix}n_\varphi[i] \Delta t\\0\\n_\psi[i]\Delta t\\0\\0\\0\end{bmatrix}}_{\substack{\text{Difference in}\\\text{angular velocity}}}
\end{align}

Assuming intended locomotion along the $z$-axis and unintended locomotion along the $x$-axis where $R$ is the sphere radius.
No movement upwards is possible since the robot moves on a flat ground, hence the entry is zero.

\subsection{Simulated Datasets}

Figure~\ref{fig:simulatedDatasets} shows a noise free trajectory through a simulated environment as well as a trajectory with noisy sensors.
In the noise-free case, the trajectory slightly bends in the direction of disturbance torque sideways while also varying in velocity in the intended direction of travel.
In the noisy case the ideal straight trajectory is assumed, hence the planes enclosing the room are sensed multiple times. 
In particular we see that the trajectory of the robot leads through one of the sensed planes. 
Also, the further the sensed points are the noisier they appear, which is consistent with the larger influence of the pose error at higher distances.

\begin{figure*}
	\centering
	\includegraphics[width=0.51\textwidth]{images/perfect_bottom}\hfill
	\includegraphics[width=0.48\textwidth]{images/noisy_pose_and_range_bottom}\\
	\includegraphics[width=0.51\textwidth]{images/perfect_top}\hfill
	\includegraphics[width=0.48\textwidth]{images/noisy_pose_and_range_top}\\
	\includegraphics[width=0.51\textwidth]{images/perfect_top_view}\hfill
	\includegraphics[width=0.48\textwidth]{images/noisy_pose_and_range_top_view}\\\vspace{0.5cm}
	\includegraphics[width=0.25\textwidth]{images/sim_frame_00.png}\hfill
	\includegraphics[width=0.25\textwidth]{images/sim_frame_02.png}\hfill
	\includegraphics[width=0.25\textwidth]{images/sim_frame_06.png}\hfill
	\includegraphics[width=0.25\textwidth]{images/sim_frame_08.png}
	\caption{Two simulated datasets of a long hallway of size $\SI{4}{\meter}\times\SI{3}{\meter}\times\SI{100}{\meter}$. Ideal noise free (left), noisy pose ($\mathcal{N}_\varphi = \mathcal{N}_\psi(\mu = 0.0001, \sigma= 0.00001)$) and noisy range measurements ($\mathcal{N}_r(\mu = 0, \sigma = 0.001)$) (right). Below: Initial sequence of the simulation of the ideal dataset showing the simulated motion of the robot.}
	\label{fig:simulatedDatasets}
\end{figure*}