### 2021-05-12 Latex/sections/acknowledgement.tex
Added acknowledgements

%!TEX root = ../main.tex
\section*{Acknowledgement}
\todo[inline]{Formulate these: ESA funding (?), elite netzwerk bayern (?), cylon floatsat group}