### 2021-05-12 Latex/sections/acknowledgement.tex
Updated conclusion and acknowledgment

%!TEX root = ../main.tex
\section*{Acknowledgement}
\todo[inline]{Formulate these: ESA funding (?), elite netzwerk bayern (?)}

We thank the Cylon team (Timothy Randolph, Ignacio Dorado, Camilo Andres Reyes) for providing the dataset and documentation of the experimental setup used in this work.