### 2021-05-13 Latex/sections/acknowledgement.tex

%!TEX root = ../main.tex
\section*{Acknowledgement}

The authors thank the Cylon team (Ignacio Dorado Llamas, Timothy Randolph, Camilo Andres Reyes Mantilla) for designing and building the prototype and Timo Burger for acquiring the floating table data set.
Special thanks to Dieter Ziegler and Sergio Montenegro for supporting our work.
We acknowledge funding from the ESA Contract No. 4000130925/20/NL/GLC for the ``DAEDALUS -- Descent And Exploration in Deep Autonomy of Lava Underground Structures'' Open Space Innovation Platform (OSIP) lunar caves-system study and the Elite Network Bavaria (ENB) for providing funds for the academic program ``Satellite Technology''.