### 2021-07-03 Latex/rebuttal.txt
Done reviewer one comments

 
> Manuscript Number: OPHOTO-D-21-00020  
> Towards Spherical Robots For Mobile Mapping In Human Made
> Environments
>
> Dear Mr. Bredenbeck,     
> 
> Thank you for submitting your manuscript to ISPRS Open Journal of
> Photogrammetry and Remote Sensing.
> 
> I have completed my evaluation of your manuscript. The reviewers
> recommend reconsideration of your manuscript following revision. I
> invite you to resubmit your manuscript after addressing the comments
> below. Please resubmit your revised manuscript by Jul 13, 2021.  
> When revising your manuscript, please consider all issues mentioned
> in the reviewers' comments carefully: please outline in a cover
> letter every change made in response to their comments and provide
> suitable rebuttals for any comments not addressed. Please note that
> your revised submission may need to be re-reviewed.     
> 
> To submit your revised manuscript, please log in as an author at
> https://www.editorialmanager.com/ophoto/, and navigate to the
> "Submissions Needing Revision" folder under the Author Main Menu.
> 
> ISPRS Open Journal of Photogrammetry and Remote Sensing values your
> contribution and I look forward to receiving your revised
> manuscript.
>   
> Kind regards,     
> 
> George Vosselman   
> Editor-in-Chief   
> ISPRS Open Journal of Photogrammetry and Remote Sensing    
>  
> Editor and Reviewer Comments:     
> 
> Editor:
> As you will see below both reviewers appreciate your work and don't
> have concerns regarding the methodology or results. Reviewers, in
> particular Reviewer 2, however, do have a large number of
> recommendations to improve the readability of your manuscript. Given
> the number of recommendations, I ask for a major revision. The
> recommendations should, however, be easy to implement.
>  
>   
> Reviewer 1: The manuscript presents the concept of a spherical
> mobile mapping system, and a data registration method utilizing the
> geometric constraints available in human-made environments,
> e.g. planar surfaces.
> 
> The manuscript has novel content and is well written, and can be
> accepted after minor revision. My comments and suggestions to be
> considered by the authors are as follows:
> 
> Lines 27-31: Spherical shell can protect the sensors inside, but
> this does not prevent the external disturbances such as dust or
> other impurity staining the sphere itself, affecting the performance
> of the system. So in principle this does not change the situation
> from using protective covers on "traditional" mobile mapping
> systems.

We clarified that the shell can enclose all sensor without requiring a 
number of points of connection as individual enlosings for all sensors
would. Further we added the requirement, that the shell is very 
durable.

> Lines 111-116: Modern kinematic scanning systems collect point cloud
> data with high speed, so also kinematic point clouds are often dense
> enough for robust plane determination. And on the other hand, sparse
> point clouds are not optimal for point-to-point matching due to the
> long distances between adjacent points.

We clarified that also the lilmited FOV plays a role in not finding 
planes for robust correspondences. 
FIXME @Andreas and @Dorit: I (Anton) disagree can you double check?

> Lines 244-246: The Livox ranging precision is given twice in this
> sentence.
> 
> Figure 4: The x-scale is given in figure caption, but I would
> suggest adding it to the figure also.

FIXME Redoing the experiment anyways

> Conclusion: Whereas the title, Abstract and Introduction concentrate
> mostly on the spherical mobile mapping, Conclusion deals almost
> entirely with the topic of registration, thus feeling somewhat
> separated from the rest of the manuscript. Reading the Conclusion,
> one could not expect that the title of the manuscript is "Towards
> Spherical Robots For Mobile Mapping In Human Made Environments".

We added more context about spherical robots and put the registration 
method into context.

> 
> Reviewer 2: The authors have written a very interesting paper on the
> topic of mobile mapping with a spherical robot in human-made
> environments. In a rather slim paper and pleasantly concise form,
> they present their approach and show first results based on
> simulated and real measurement data. The reading flow is mostly
> pleasant and no irrelevant information is given. Conversely, the
> authors leave out important details in several places, which would
> be very helpful for the reader's understanding. In the following,
> there are several points for improvement and critical questions that
> should be taken into account by the authors. Due to the large number
> and the importance of central points, I recommend a major revision.
> 
> - A short outline at the end of the introduction is missing and
> would give the reader a better overview of the further content and
> should therefore be added.
> 
> - In the entire article, no reference is made to the targeted or
> required accuracy potential with regard to the quality of the pose
> determination and the final point cloud, which in my view requires
> at least a rough classification.
> 
> - For the short state of the art of existing mobile systems, which
> is mainly based on the authors' own systems, in my opinion, at least
> the NavVis VLX system could be added, which in comparison is newer
> and as a human-operated system for mapping is one of the benchmarks.
> 
> - The authors mention unknown and dangerous environments as a
> possible field of application, in which the advantages of the
> spherical system should come into play. However, only one space
> application is mentioned as a specific example, which may be valid
> in itself, but does not fit the topic of the paper very
> well. Application scenarios should therefore be supplemented by
> terrestrial areas and in particular by human-made environments,
> which better motivate the future use and justification of such
> systems for the reader.
> 
> - The advantages of a spherical mobile platforms are discussed, but
> nothing is said about possible disadvantages or
> limitations. Moreover, I wonder whether the "unique advantages"
> mentioned in the abstract are really so remarkable. For example,
> what about the recording constellation, which in my view is
> relatively unfavourable on the ground and leads to restricted fields
> of view and weak angles of incidence when mapping the
> ground. Another advantage stated by the authors is that the shell
> protects the sensors, which is reasonable. However, this protection
> is only necessary because you are moving directly on the ground. In
> addition, can dust and dirt accumulate on the outer shell and
> influence the acquisition?
> 
> - Section 2.1 discusses existing laser scanner-based SLAM
> approaches, concluding with the statement (page 2, column 1, line
> 30-33) that all of these approaches do not include the structural
> properties of human-made environments in the calculations. This
> statement is not true. There are a variety of SLAM approaches
> (e.g. Jung et al. 2015; Nguyen et al.  2006), but also other
> applications with recursive state estimation, which make use of this
> additional information in the context of georeferencing. This should
> be investigated and reworked in more detail.
> Jung, J., Yoon, S., Ju, S., and Heo, J. (2015). Development of Kinematic
> 3D Laser Scanning System for Indoor Mapping and As-Built BIM Using
> Constrained SLAM. In: Sensors (Basel, Switzerland), 15(10), DOI:
> https://doi.org/10.3390/s151026430
> Nguyen, V., Harati, A., Martinelli, A., Siegwart, R., and Tomatis, N.
> (2006). Orthogonal SLAM: A Step Toward Lightweight Indoor Autonomous
> Navigation. In: IEEE/RSJ International Conference on Intelligent Robots
> and Systems, DOI: https://doi.org/10.1109/IROS.2006.282527
> 
> - The introduction in Section 3 should be revised as it is not very
> intuitive for the reader and something about the objectives and
> motivation for this section is missing. However, this is also
> directly related to the general lack of outline beforehand.
> 
> - In section 3, the authors introduce their own algorithm for
> registration, but they do not give any further information about the
> temporal context in which this takes place. This is probably done in
> the context of each epoch, although it is also not clear how an
> epoch is defined. Is it a single full rotation of the laser scanner
> or that of the sphere? And does the registration always take place
> only locally in relation from the current epoch to the last epoch or
> with a window approach or even globally?
> 
> - Section 3.2 deals with the extraction of planes, but the following
> questions remain unclear: Are the planes always searched for only in
> the context of the observational data of the current epoch or with
> regard to the already registered global point clouds? How many
> planes are needed at all and what demands are made on their spatial
> distribution? The abstract mentioned parallel and perpendicular
> planes, but this is not discussed further here. Are only these used
> and if so, why? Furthermore, what about planes that are not
> necessarily parallel or perpendicular (e.g. doors, cupboards)? How
> is the flatness of a plane checked? How are outliers dealt with,
> especially with regard to existing measurement and motion noise?
> 
> - Throughout Section 3, numerous variables are not introduced and
> are not explained in more detail, so that a more precise
> interpretation of the individual equations is unfortunately very
> difficult.
> 
> - What orders of magnitude were finally used for the range noise
> model in Section 4.2 and why was this not also done for the angular
> resolution?
> 
> - In the context of the real experiment in Section 5, a major
> simplification is made. Even if this simplification is certainly
> suitable for the beginning, this special case comes quite surprising
> here...especially in comparison to the simulation and the previously
> described topic. Have other measurements not yet been carried out or
> what is the background? At least this simplification should be
> mentioned beforehand, otherwise the reader will have false
> expectations.
> 
> - Wouldn't it also be interesting to compare the estimated pose
> parameters with the nominal values from the simulation within the
> framework of Section 6.1? So far, the focus has only been on the
> point cloud.
> 
> - Why is there no qualitative validation for the results of the real
> data in Section 6.2? A reference scan based on a terrestrial laser
> scanner would be a possibility or otherwise based on the flatness of
> known planes in the registered point cloud?
> 
> - In the Conclusion, the parallelism and perpendicularity within the
> environment is mentioned again, although this did not play a role in
> the main part of the paper. Therefore, the question arises why do
> the planes have to be parallel or perpendicular at all and why is
> this mentioned at the beginning? This is not clear, since no
> additional constraints on this additional information have been
> introduced so far. There are also already other methods available
> that take this idea of geometric constraints for georeferencing into
> account.
> 
> Further open questions/comments are:
> Page 1, Line 15: Double use of "protect", which should be replaced
> by another word.

We removed first instance.

> Page 1, Line 21: The link for the "second option" to the use of an
> IMU is not necessarily obvious.

We added IMU as sensor for coarse estimation.

> Page 1, Column 2, Line 52-56: Complicated sentence which should be
> reformulated.

We have split into two sentences

> Page 2, Column 2, Line 15: The term "long stop" is not to be
> interpreted clearly in this context. How is "long" defined in this
> sense in terms of time or space?

The specified stop is required to find enough points for plane
detection.

> Page 3, Column 1, Line 6-8: A specific reference should be given for
> the first half of the sentence, as reference 19 probably only refers
> to the second part.

FIXME! Added todo for Dorit/Andreas as they wrote this

> Page 3, Column 1, Line 23-26: Not all variables were introduced
> accurately

We have introduced missing variables and clarified rho.

> Page 3, Column 1, Line 33-34: Selection of references for this?

We have added plane segmentation sources.

> Page 3, Column 1, Line 34-35: But what demands are made on the
> assignment? This is unclear to the reader, also in connection with
> the general quality requirements of the whole approach.

We have added a specification for "good enough for plane detection".

> Page 3, Column 1, Line 38: How can the threshold value be specified,
> since it is probably not identical for all applications and
> environments? Later, there is also no indication of the choice of
> the value in the own simulated and real data sets.

FIXME: Added todo for Fabi

> Page 3, Column 1, Line 50: What does the index k describe and what
> is the difference to the index i from Section 3.2?

We have clarified indeces k=planes i=points.

> Page 3, Equation (1): Again, some variables are not specified. For
> example, the source of tx ty tz is unclear.

We have introduced tx ty tz properly.

> Page 3, Equation (3): Equation unclear regarding the use of the 2 in
> the exponent as well as the index.

We have specified that we're using the square of the L2 norm.

> Page 3, Equations (4)-(5): Here I don't understand how to get to the
> elements of the normal vector through the transformation. Moreover,
> these individual elements were not introduced in advance.

We have corrected indeces and introduced the elements.

> Page 3, Equation (9): The variable alpha is not described and
> explained

An explaination of alpha was added.

> Page 4, Equation (10): Description of the individual variables is
> missing and, moreover, rho has already been used as the distance
> parameter of the Hesse-normal form

we have replaced rho with zeta and introduced the variables.

> Page 4, Equation (12): Is epsilon here the same variable as the
> threshold in Section 3.3? Probably not.

We have renamed epsilon and clarified the difference to sec 3.3.

> Page 4, Column 2, Line 15: "some dimensions": Which ones and why
> exactly are they chosen? Are they the same ones all the time?

FIXME: chose better example, added todo for Fabi

> Page 4, Column 2, Line 20-25: Reference is missing

We haded a citation.

> Page 4, Column 2, Line 55: Doubled information regarding the
> distance measurement precision

We removed duplicated information.

> Page 5, Column 1, Equation (18): Delta t is not introduced and there
> is no indication of how large the time steps are in the simulation.

We have introduced delta t and give a default value.

> Page 5, Column 2, Line 14: Should it be "out" or "our"?

We have corrected this.

> Page 5, Column 2, Line 27-28: What are the specifications of the
> used IMU with regard to its uncertainty budget?

We have added precision of IMU and added the source.

> Page 6, Figure 1: How are these values to be interpreted and from
> where are they derived? This is unclear.

We have specified this. It is chosen to still allow plane detection.

> Page 8, Column 1, Table 1: What is missing is a classification of
> the values in relation to the expectations / demands...which are,
> however, still missing in advance

FIXME: -> I don't understand this comment? TODO

> Page 8, Column 1, Line 37-43: At this point, an interpretation of
> the significant second peak in the right-hand diagram in Figure 4
> should be made. This has not been discussed so far.

FIXME: -> added todo REDO experiment

> Page 8, Column 1, Line 51: "ten scans" means ten individual scan
> lines or 10 epochs or is this identical? What kind of temporal
> offset is there in between and wouldn't a motion model have to be
> taken into account for this or is an identical pose assumed for the
> ten scans?

We have specified which scans and that we assuem constant pose.

> Page 8, Column 1, Line 55-56: What could be the cause of this?

We have specified the outlier dependence.

> Page 8, Column 2, Line 44: This sentence always applies and should
> not remain here.

We have removed this sentence.

> Page 9, Figure 4: The two point clouds in the middle of figure 4 are
> difficult to interpret in terms of their orientation. Could it be an
> option to add a coordinate system?

It was specified in the figure label that lateral images have the same
orientation.