### 2021-07-07 Latex/sections/realData.tex
LPC results added, rolling sphere added.

%!TEX root = ../main.tex
\section{Experimental Setup}\label{sec:experimentalSetup}

We also tested our algorithm on two real-world datasets that adhere to different motion profiles and laser scanner types with different scanning patterns.

The first dataset is collected by a line scanner, in particular a SICK LMS141 industrial scanner, inside a sphere of a diameter of \SI{20}{\centi\meter} that lies on a floating desk, which is air-pressurized, such that the sphere is hovering.
On this floating desk, the sphere can rotate freely about all axis while being fixed in position. 
Hence, in this experiment the optimization space can be reduced to a rotation.
Further, without the motion, the 2D laser cannot obtain a 3D model from the environment, thus requiring registration. 
The sphere is equipped with a low-cost IMU, i.e., a PhidgetSpatial Precision 3/3/3, to estimate the orientation, a battery pack and a raspberry pie for processing.
This IMU provides a precision of \SI{76.3}{\micro g} in linear acceleration, \SI{0.02}{\degree\per\second} in angular velocity about the x and y axis, and \SI{0.013}{\degree\per\second} in angular velocity about the z axis \cite{PhidgetSpatial}.  
Figure~\ref{fig:float-sat-sphere} shows the experimental setup. The left column of figure~\ref{fig:cylon-corrected} shows the pre-registered resulting point cloud. 
We see that the point cloud is rather noisy, and in particular, the walls are sensed multiple times, hence appearing very blurry.

\begin{figure}
	\centering
	\begin{minipage}[c]{0.45\textwidth}
		\centering
		\includegraphics[width=0.33\textwidth, height=0.33\textwidth]{./images/sphere-frame-1.eps}\hfill
		\includegraphics[width=0.33\textwidth, height=0.33\textwidth]{./images/sphere-frame-2.eps}\hfill
		\includegraphics[width=0.33\textwidth, height=0.33\textwidth]{./images/sphere-frame-3.eps}\\
		\includegraphics[width=0.33\textwidth, height=0.33\textwidth]{./images/sphere-frame-4.eps}\hfill
		\includegraphics[width=0.33\textwidth, height=0.33\textwidth]{./images/sphere-frame-5.eps}\hfill
		\includegraphics[width=0.33\textwidth, height=0.33\textwidth]{./images/sphere-frame-6.eps}\hfill
	\end{minipage}\\
	\vspace{0.1cm}
	\begin{minipage}[c]{0.45\textwidth}
		\centering
		\includegraphics[width=0.495\textwidth]{./images/cylon_exploded}\hfill
		\includegraphics[width=0.495\textwidth]{./images/cylon_setup}\hfill
	\end{minipage}
	\caption{Data acquisition using floating sphere. Top: A sequence of orientations the sphere assumes during data acquisition. Bottom: Hardware setup of the sphere.} 
	\label{fig:float-sat-sphere}
\end{figure}

The second datasets is collected by a 3D scanner, namely the LIVOX MID 100, which is the same sensor that the simulation uses.
We put the laserscanner into an acrylic glass spherical shell together with three IMUs and roll it by hand with a slow, constant motion in a home indoor environment.
Thus, this time almost all six degrees of freedom are used for optimization.
We exlude a translation in the global up- or downwards direction, though, since we also locked that dimension when collecting the data.
Figure~\ref{fig:prototype} shows the prototpye.
The left column of figure~\ref{fig:jasperhome} shows the pre-registered point cloud, applying only the pure transformation found by the recorded pose.
Although the robot has been moved carefully slow, you see that pose error accumulates in such a way that the same wall is visible multiple times.

\begin{figure}
  \includegraphics[width=0.49\linewidth]{./images/photoReal1.jpeg}
  \includegraphics[width=0.50\linewidth]{./images/photoReal3.jpeg}
  \caption{Prototype used for data acquisition with the rolling sphere. The main payload is the Livox Mid-100 laser-scanner. For pose-estimation, three IMUs of the manufacturer Phidget are placed inside and a Raspberry Pi 4 for the calculations. On the top are two batteries, and on the bottom one voltage stabilizer and breakout box of the laser-scanner. }
  \label{fig:prototype}
\end{figure}

\iffalse % This is commented out because we couldn't produce results
	\subsection{RADLER}

	The second dataset is recorded using the RADLER system as described in~\cite{Borrmann2020-RADLER}.
	While not a spherical robot, this system shows many similarities with respect to its motion with a spherical robot.
	The system was manually steered along with a squared set of hallways that loop back to the initial starting point inside the old mathematics building at the University of Würzburg (denoted ``The Circle'' in~\cite{Borrmann2020-RADLER}).

	Figure~\ref{fig:radler-mathe} shows the system setup as well as the resulting point cloud.
	We see that with continuing measurement time, the drift increases, and the hallways are senses multiple times with a clear offset to each other.

	\begin{figure*}
		\centering
		\begin{minipage}[c]{\textwidth}
			\includegraphics[width=0.2925\textwidth]{./images/radler_setup}\hfill
			\includegraphics[width=0.7\textwidth]{./images/mathe_top_view}
		\end{minipage}
		\caption{Data acquisition using the RADLER system. Left: System setup (courtesy of Borrmann~\cite{Borrmann2020-RADLER}). Right: Unprocessed resulting point cloud.}
		\label{fig:radler-mathe}
	\end{figure*}

	\subsection{Descend Sphere REMOVE THIS SECTION}

	This was recorded using a LIVOX-Mid 100 laser scanner mounted inside a transparent shell. 
	The robot contained three Phidget 3/3/3 Spatial IMUs and was attached to a \SI{50}{\meter} tether cable that was rolled on a spool which in turn was connected to an angular encoder. 
	From these sensors, we obtained orientation information as well as height information. 
	The setup was then brought into a high indoor building (\href{https://www.sfs-w.de/feuerwehrschule/virtueller-rundgang.html}{Firefighter School in Würzburg}) where the system was manually descended from a balcony scanning the building interior.
	The robot covered a distance of approximately \SI{22}{\meter} in about \SI{402}{\second}.
	Additionally, the building was scanned using a terrestrial laser scanner to provide a reference.
	The measurements were then pre-registered before applying our plane-based correction. 
	Figure~\ref{fig:experimental-setup} shows the acquired 3D point clouds and the experimental setup.

	\begin{figure}
		\centering
		\begin{minipage}[c]{0.25\textwidth}
			\includegraphics[width=\textwidth]{images/scan-riegl-fire}\\
			\includegraphics[width=\textwidth]{images/scan-sphere-fire}
	  	\end{minipage}\hfill
	  	\begin{minipage}[c]{0.2075\textwidth}
	  		\includegraphics[width=\textwidth]{./images/lidarsetup}
	  	\end{minipage}
		\caption{Reference 3D point cloud acquired with a RIEGL 3D terrestrial laser scanner (top left). The 3D point cloud acquired by the test sphere after the pre-registration (bottom left) and a picture of the test setup (right).}
		\label{fig:experimental-setup}
	\end{figure}
\fi